---
title: "Real data example"
author: "Han-Yueh Lee"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Real data example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Load data  
```{r}
data("NCEP_U")
data("NCEPlatlon")

dim(NCEP_U)
dim(NCEPlatlon)
```

# Preprocessing  
## Centering  
```{r}
y_mat_all = NCEP_U
s_new = NCEPlatlon
s_new_id = data.frame(cbind(id=1:nrow(s_new), 
                            lon=s_new[,1], 
                            lat=s_new[,2]))

y_mat_all_c = y_mat_all - 
        matrix(rep(colMeans(y_mat_all), nrow(y_mat_all)),
               nrow=nrow(y_mat_all),ncol=ncol(y_mat_all),
               byrow=T)
y_mat_all_c = as.matrix(y_mat_all_c)
```

## Training & testing set  
```{r}
s = expand.grid(unique(s_new[,1])[seq(1,17,by=2)],
                unique(s_new[,2])[seq(1,17,by=2)])
colnames(s)=c("lon","lat")
s = data.frame(s)
s_id = dplyr::left_join(s, s_new_id, by=c("lon", "lat"))
train_index = s_id$id

# transform format
s_new = as.matrix(s_new)
s = as.matrix(s)
y_mat = y_mat_all_c[1:360,train_index]
y_mat_new = y_mat_all_c[361:480,train_index]
```

## Long format table  
```{r}
#location table
location = data.frame(s_new)
colnames(location)=c("lat","lon")
location$set = rep("test", nrow(s_new))
location$set[train_index] = "train"
location$id = 1:nrow(location)
y_test = y_mat_all_c[361:480,]
test_idx = which(location$set=="test")

#long-format space-time table
library(magrittr)
space_wide = data.frame(y_mat_all_c)
space_wide$time = 1:nrow(space_wide)
data_long = space_wide %>%
        tidyr::gather(id, z, -time) %>%
        dplyr::mutate(id= sub("V","",id)) %>%
        dplyr::mutate(training_set=(time %in% c(1:360) & id %in% train_index)) %>%
        dplyr::mutate(id = as.numeric(id)) %>%
        dplyr::left_join(s_new_id, by="id")
```

# Tuning hyper-parameters  
## Candidate sets
```{r}
tau_cv = 2^seq(1,20,by=1)
k = 1:40
```

## Tuning process
```{r, eval=FALSE}
cv_data = temp_spat_cv_final3(s=s, y_mat=y_mat,
                         sigma2_eps = 1,
                         itermax = 30, 
                         tol=0.001, 
                         tau=tau_cv, k=k)
cv_data$best_list$k # = 39
cv_data$best_list$stau # = 32768
```

# Fitting model  
```{r}
temp_spat_result = temp_spat_data_em3(s = s, 
                                         s_new = s_new,
                                         y_mat = y_mat,
                                         y_mat_new = y_mat_new,
                                         k = 39,
                                         tau = 32768,
                                         sigma2_eps=1, 
                                         itermax = 30, 
                                         tol = 10^(-3))
```

# Spatial prediction and temporal forecast (point prediction)  

1. Given arbitrary time t, output spatial prediction surface at time t  
2. Given arbitrary time t-1, output temporal forecast surface at time t   

## t = 361, 420, 480  
```{r, fig.width=8,fig.height=6}
t_set = c(361, 420, 480)
library(lattice)
library(gridExtra)
plist=list()
for(i in 1:3){
        t = t_set[i]
        y_comp = data.frame(observe = y_mat_all_c[t,],
                            pred = temp_spat_result$pred_result$y_spat[t-360,],
                            forecast = temp_spat_result$pred_result$y_temp[t-360,],
                            lat=s_new[,1],
                            lon=s_new[,2])
        plist[[1+(i-1)]] = levelplot(observe~lon*lat, 
                       data=y_comp,
                       scales = list(draw = FALSE), xlab = NULL, ylab = NULL, 
                       at=seq(min(y_comp[,1:3]), max(y_comp[,1:3]), length.out=100),
                       col.regions = heat.colors(100)[length(heat.colors(100)):1], 
                       main=paste0("t",t,"_true"))
        plist[[4+(i-1)]] = levelplot(pred~lon*lat, 
                       data=y_comp,
                       scales = list(draw = FALSE), xlab = NULL, ylab = NULL, 
                       at=seq(min(y_comp[,1:3]), max(y_comp[,1:3]), length.out=100),
                       col.regions = heat.colors(100)[length(heat.colors(100)):1], 
                       main=paste0("t",t,"_pred"))
        plist[[7+(i-1)]] = levelplot(forecast~lon*lat, 
                       data=y_comp,
                       scales = list(draw = FALSE), xlab = NULL, ylab = NULL, 
                       at=seq(min(y_comp[,1:3]), max(y_comp[,1:3]), length.out=100),
                       col.regions = heat.colors(100)[length(heat.colors(100)):1], 
                       main=paste0("t",t,"_forecast"))
}
grid.arrange(grobs=plist, ncol=3)
```

## Empirical coverage rate of spatial prediction  
```{r}
y_observe = y_mat_all_c[361:480,]
mean(y_observe <= temp_spat_result$pred_result$y_spat_up &
             y_observe >= temp_spat_result$pred_result$y_spat_lw)
```

## Empirical coverage rate of spatial prediction  
```{r}
mean(y_observe <= temp_spat_result$pred_result$y_temp_up &
             y_observe >= temp_spat_result$pred_result$y_temp_lw)
```


# Performance  
## $MSE_{spat}$  
```{r}
sum((temp_spat_result$pred_result$y_spat[,test_idx]-y_mat_all_c[361:480,test_idx])^2)/(120*289)
```

## $MSE_{temp}$  
```{r}
sum((temp_spat_result$pred_result$y_temp[,test_idx]-y_mat_all_c[361:480,test_idx])^2)/(120*289)
```

# Compare to Model II  
## Generate $\Phi$ matrix  
```{r}
library(FRK)
library(sp)
G1 = auto_basis(data=s_new %>% SpatialPoints(),
               nres=1,type="Gaussian")
show_basis(G1)
phi_res1 = as.matrix(eval_basis(G1, as.matrix(s)))
phi_res1_new = as.matrix(eval_basis(G1, as.matrix(s_new)))
```

## Fitting Model II
```{r}
model2 = temp_spat_data_em_model2(s=s, s_new=s_new, 
                                 y_mat=y_mat, y_mat_new=y_mat_new,
                                 bas_mat=phi_res1, bas_mat_new =phi_res1_new ,
                                 k=ncol(phi_res1), 
                                 sigma2_eps=1, itermax = 30, tol = 0.001)
```

## Performance  
## $MSE_{spat}$  
```{r}
sum((model2$pred_result$y_spat[,test_idx]-y_mat_all_c[361:480,test_idx])^2)/(120*289)
```

## $MSE_{temp}$  
```{r}
sum((model2$pred_result$y_temp[,test_idx]-y_mat_all_c[361:480,test_idx])^2)/(120*289)
```

