---
title: "Simulation Example"
author: "Han-Yueh Lee"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Simulation Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Generate simulation data (snr = 0.25)

## overall setting
```{r, fig.show='hold'}
# time length
n_train=100
n_test=50
n=n_train+n_test

# number of location
p=50
s <- matrix(seq(-5, 5, length = p),nrow=p,ncol=1)
s_new <- matrix(seq(-5, 5, length=3*p), nrow=3*p, ncol=1)
# s = matrix(seq(-5,5,by=0.2))
# s_new = matrix(seq(-5,5,by=0.1))
# p = nrow(s)


# dim of eigen-space
k=2

# phi, A, and Lambda matrix
phi_1 <- exp(-s^2)/norm(as.matrix(exp(-s^2)),type="F")
phi_2 <- s*exp(-s^2)/norm(as.matrix(s*exp(-s^2)), "F")
phi_mat <- cbind(phi_1, phi_2)
phi_mat_new <- new_loc_phi(s_new, s, B=B_gen(s, phi_mat))
A = matrix(c(0, 0,
             0, 0.8), nrow=2, byrow=T)
lambda_mat=matrix(c(9,0,
                    0,9),nrow=2, byrow=T)

# plot phi
for(i in 1:2){
        plot(s, phi_mat[,i], type="l", 
             xlab="s",ylab="loading", col="red", lwd=2)
}
```

## signal to noise ratio (snr) = 0.25  
```{r}
snr = 0.25
sigma_sq = sum(diag(lambda_mat))/(snr*p)
V = V_mat_calc(A, lambda_mat)
```

## generate simulation data
```{r}
itr = 1
set.seed(itr)
gen_result1 <- data_gen_1d_2(s_old_mat=s, s_new_mat=s_new, k=2, n=n,
                             lambda_mat=lambda_mat,
                             sigma_sq=sigma_sq,A=A, V=V)

y_mat <- gen_result1$y_mat[1:n_train,]
y_mat_new <- gen_result1$y_mat[(n_train+1):n,]
                
xsi_mat = gen_result1$xsi_mat[1:n_train,]
new_xsi_mat = gen_result1$xsi_mat[(n_train+1):n,]

signal_mat = xsi_mat %*% t(phi_mat)
signal_mat_new = new_xsi_mat %*% t(phi_mat_new)
```

## take a look  
### dim check  
```{r}
dim(y_mat) # 100(time) by 50(location) matrix
dim(xsi_mat) # 100(time) by 2(states) matrix
dim(signal_mat) # 100(time) by 50(location) matrix
```

### observation vs signal  
```{r, fig.width=6, fig.height=4}
par(mfrow=c(2,3))
for(t in c(1,25,50,75,100)){
        plot(s, y_mat[t,],type="l",lwd=2,main=paste0("t = ", t),ylab="Y_t(s)")
        lines(s, signal_mat[t,],lwd=2,col="red")
}
```

### time-series of each state  
```{r, fig.show='hold'}
ts.plot(xsi_mat[,1],ylab="",main="xi_1")
ts.plot(xsi_mat[,2],ylab="",main="xi_2")
```

# Tuning hyper-parameters  

## Hyper-parameter set  
```{r}
tau_cv = 2^seq(-10,10,by=1)
k_select=c(1,2,3,4)
```


```{r eval=FALSE}
# parallel computing setting  
library(foreach)
library(doParallel)
cl = makeCluster(6)
registerDoParallel(cl)

# tuning
cv_result_all = temp_spat_cv_final2(s=s, y_mat=y_mat,sigma2_eps = 1,
                                    itermax = 30, tol=0.1, tau=tau_cv, k=k_select)
cv_result_all$best_list$k # = 2
cv_result_all$best_list$stau # = 8
```

Since the tuning process is too time-consuming, we load the tuning results.  
```{r}
data("cv_result_all")
cv_result_all$best_list$k
cv_result_all$best_list$stau
```

# Fitting model  
## Parameter estimation  
```{r}
EM_result <- EM_func3(s=s, y_mat=y_mat, 
                      y_mat_new=y_mat_new, 
                      sigma2_eps=1, 
                      itermax = 30, 
                      tol = 0.001, 
                      tau=8, 
                      k=2)
```

### Check estimation of Phi  
```{r, fig.show='hold'}
plot(s, phi_mat[,1],type="l",col="gray",lwd=2,
     ylab="loading",ylim=c(-0.4,0.4))
lines(s, EM_result$Phi[,1],col="red",lwd=2)

plot(s, phi_mat[,2],type="l",col="gray",lwd=2,
     ylab="loading",ylim=c(-0.4,0.4))
lines(s, EM_result$Phi[,2],col="red",lwd=2)
```

### Check estimation of measurement error  
```{r}
cat(paste0("True variance of measurement error is ",sigma_sq,"\n",
           "Estimated variance of measurement error is ",EM_result$sigma2_eps))
```

### Check estimation of A  
```{r}
A #true A  
EM_result$A #estimated A
```

### Check Ca  
```{r}
V # true covariance matrix of a_t
EM_result$Ca # Estimated covariance matrix of a_t
```

## kalman filter, spatial prediction, & temporal forecast  
```{r}
new_phi <- new_loc_phi(s_new=s_new, 
                               s_old = s, 
                               B = B_gen(s, EM_result$Phi))
pred_result = pred_func(new_phi=new_phi, 
                        y_mat_new=y_mat_new, 
                        EM_result=EM_result)
```

### check new_phi  
```{r, fig.show='hold'}
plot(s_new, pred_result$new_phi[,1],type="l",
     col="red",lwd=2,ylab="loading",ylim=c(-0.4,0.4))
plot(s_new, pred_result$new_phi[,2],type="l",col="red",
     lwd=2,ylab="loading",ylim=c(-0.4,0.4))
```


### check xi result (filter & prediction)  
```{r, fig.width=5,fig.height=5}
par(mfrow=c(2,2))
plot(1:n_test,pred_result$xi_result$xi_pred[1,],type="l",
     xlab="time",ylab="xi_1")
plot(1:n_test,pred_result$xi_result$xi_pred[2,],type="l",
     xlab="time",ylab="xi_2")

acf(pred_result$xi_result$xi_pred[1,],main="xi_1_series")
acf(pred_result$xi_result$xi_pred[2,],main="xi_2_series")
```


## check spatial prediction (point)  

### given arbitrary time t, predict spatial surface    
```{r, fig.width=8,fig.height=6}
par(mfrow=c(2,2))
for(t in c(20, 30, 40, 50)){
        signal = signal_mat_new[t,]
        observation = signal + rnorm(n = length(signal), sd = sqrt(sigma_sq))
        pred = pred_result$y_spat[t,]
        pred_up = pred_result$y_spat_up[t,]
        pred_lw = pred_result$y_spat_lw[t,]
        plot(observation, type="l", col="gray",
             xlab="location", ylab="response_measure",
             main=paste0("time = ",t),ylim=c(-4,4))
        lines(signal,col="blue")
        lines(pred, col="red")
        lines(pred_up, col="red",lty=2)
        lines(pred_lw, col="red",lty=2)
        legend("bottomright",
               legend = c("observation","signal","predict","interval"),
               col=c("gray","blue","red","red"),
               lty=c(1,1,1,2),cex=0.5)
} 
```

### empirical coverage rate of spatial prediction   
```{r}
set.seed(1)
observation = signal_mat_new + rnorm(n=length(signal_mat_new),
                                     mean=0,sd = sqrt(sigma_sq))
mean(observation<=pred_result$y_spat_up & 
             observation>=pred_result$y_spat_lw)
```


## temporal forecast  
### given arbitrary time t, forecast spatial surface in time (t+1)   
```{r, fig.width=8,fig.height=6}
par(mfrow=c(2,2))
for(t in c(20, 30, 40, 50)){
        signal = signal_mat_new[t,]
        observation = signal + rnorm(n = length(signal), sd = sqrt(sigma_sq))
        forecast = pred_result$y_temp[t,]
        forecast_up = pred_result$y_temp_up[t,]
        forecast_lw = pred_result$y_temp_lw[t,]
        plot(observation, type="l", col="gray",
             xlab="location", ylab="response_measure",
             main=paste0("time = ",t),ylim=c(-4,4))
        lines(signal,col="blue")
        lines(forecast, col="red")
        lines(forecast_up, col="red",lty=2)
        lines(forecast_lw, col="red",lty=2)
        legend("bottomright",
               legend = c("observation","signal","forecast","interval"),
               col=c("gray","blue","red","red"),
               lty=c(1,1,1,2),cex=0.5)
} 
```

### empirical coverage rate of temporal forecast   
```{r}
set.seed(1)
observation = signal_mat_new + rnorm(n=length(signal_mat_new),
                                     mean=0,sd = sqrt(sigma_sq))
mean(observation<=pred_result$y_temp_up & 
             observation>=pred_result$y_temp_lw)
```

## performance measure  

### true value preparation  
```{r}
y_new_true = signal_mat_new
lag0_cov_true = phi_mat_new %*% lambda_mat %*% t(phi_mat_new)
lag1_cov_true = phi_mat_new %*% A %*% lambda_mat %*% t(phi_mat_new)
```

### performance calculation  

```{r}
performance = performance_func(y_new_true = signal_mat_new,
                               y_spat = pred_result$y_spat,
                               y_temp = pred_result$y_temp,
                               lag0_cov_true = lag0_cov_true,
                               lag0_cov_est = pred_result$lag0_cov_est,
                               lag1_cov_true = lag1_cov_true,
                               lag1_cov_est = pred_result$lag1_cov_est)
performance
```

## all in one function  
```{r}
sim_result = temp_spat_sim_em3(s=s, y_mat=y_mat, 
                      y_mat_new=y_mat_new, 
                      sigma2_eps=1, 
                      itermax = 30, 
                      tol = 0.001, 
                      tau=8, 
                      k=2,
                      s_new = s_new, 
                      y_new_true = y_new_true)
```

results:  
```{r}
sim_result$performance
```


# Compare to other models  
## Fitting model 1  
```{r}
model1_result = model1_func(s, s_new, y_mat,y_mat_new,
                            xsi_mat, phi_mat,
                            new_xsi_mat, phi_mat_new, 
                            lambda_mat, A)
model1_result$performance
```

## Fitting model 2  

### generate basis function  
```{r}
scale_set = 1.5
kkk = 10
bas_mat = auto_gau_basis_1d(s_mat = s, 
                            num_center = kkk,
                            scale = scale_set)
bas_mat_new = auto_gau_basis_1d(s_mat = s_new, 
                                num_center = kkk,
                                scale = scale_set)

```


### fitting  
```{r}
model2_result = temp_spat_sim_em_model2(s=s, s_new=s_new, 
                                 y_mat=y_mat, y_mat_new=y_mat_new,
                                 bas_mat=bas_mat, 
                                 bas_mat_new = bas_mat_new,
                                 new_xsi_mat=new_xsi_mat, 
                                 phi_mat_new=phi_mat_new,
                                 lambda_mat=lambda_mat, 
                                 A=A, 
                                 k=kkk, 
                                 sigma2_eps=1, 
                                 itermax = 30, 
                                 tol = 0.001)
model2_result$performance
```

