---
title: "Simulation Example"
author: "Han-Yueh Lee"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Simulation Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Vignettes are long form documentation commonly included in packages. Because they are part of the distribution of the package, they need to be as compact as possible. The `html_vignette` output type provides a custom style sheet (and tweaks some options) to ensure that the resulting html is as small as possible. The `html_vignette` format:

- Never uses retina figures
- Has a smaller default figure size
- Uses a custom CSS stylesheet instead of the default Twitter Bootstrap style

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))


# Generate simulation data  

## overall setting
```{r}
# time length
n_train=100
n_test=50
n=n_train+n_test

# number of location
p=50
s <- matrix(seq(-5, 5, length = p),nrow=p,ncol=1)
s_new <- matrix(seq(-5, 5, length=3*p), nrow=3*p, ncol=1)

# dim of eigen-space
k=2

# phi, A, and Lambda matrix
phi_1 <- exp(-s^2)/norm(as.matrix(exp(-s^2)),type="F")
phi_2 <- s*exp(-s^2)/norm(as.matrix(s*exp(-s^2)), "F")
phi_mat <- cbind(phi_1, phi_2)
phi_mat_new <- new_loc_phi(s_new, s, B=B_gen(s, phi_mat))
A = matrix(c(0, 0,
             0, 0.8), nrow=2, byrow=T)
lambda_mat=matrix(c(9,0,
                    0,9),nrow=2, byrow=T)
```

## snr = 0.25  
```{r}
snr = 0.25
sigma_sq = sum(diag(lambda_mat))/(snr*p)
V = V_mat_calc(A, lambda_mat)

itr = 1
set.seed(itr)
gen_result1 <- data_gen_1d_2(s_old_mat=s, s_new_mat=s_new, k=2, n=n,
                             lambda_mat=lambda_mat,
                             sigma_sq=sigma_sq,A=A, V=V)

y_mat <- gen_result1$y_mat[1:n_train,]
y_mat_new <- gen_result1$y_mat[(n_train+1):n,]
                
xsi_mat = gen_result1$xsi_mat[1:n_train,]
new_xsi_mat = gen_result1$xsi_mat[(n_train+1):n,]
```